Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-01-21 2_SSL_EES/article/article_main.tex
updated main article

where $E_i$ denotes the total energy of light at the color corresponding to the down-converted photon wavelength (Blue, Red, Orange, Yellow, Green)\cite{schubert2018light}\cite{tsao2010solid}.

@@

where $K$ is the luminous efficacy of radiation of the light source, can be computed from the spectrum and the luminosity function, which describes the sensitivity of the human eye.$K_{max}$ is the maximum luminous efficacy of radiation of a perfect light source with the same color rendering performance and correlated color temperature as the light source in question\cite{schubert2018light}\cite{tsao2010solid}.

@@

\section{Methods and Data Collection}

@@

\begin{figure*}[h!]
 \centering
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2003.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2010.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2016.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2020.pdf}
 \caption{Waterfall diagrams of the loss channels in a generic mid/high-power LED package for 2003, 2010, 2016 and 2020 (top to bottom panels), normalized to 1 Watt of electric power input (yellow bar on the left). Sub-efficiencies corresponding to each loss channel are listed below each column and described in Section \ref{sec:device_performance_metrics}. Numbers for each loss channel indicate energy losses both in relative terms of input power (in percent) at the point of the channel and absolute values (in Watts). Percentages for red, green and blue loss channels indicate losses of remaining red/green/blue light energy. The corresponding LED architectures from Figure \ref{fgr:chip_architecture_overview} are shown for reference: 2003 (top panel) - flip-chip with YGAG phosphor; 2010 (bottom panel) - flip-chip with 258 phosphor; 2016 (top panel) - flip-chip with PFS phosphor; 2020 (bottom panel) - flip-chip with SALON phosphor. Details on phosphors are provided in Table \ref{tab:phosphors} and SI section XX. Overall LED package efficiency is $\eta_L = 5.8\%$ in 2003 and $\eta_L = 12.7\%$ in 2010. Abbreviations: Scat. = Scattering. Source: own elaboration based on data from sources listed in SI, section XX.}
 \label{fgr:waterfall}
\end{figure*}

@@

        1982 & 1996 & 2015 & Patterned sapphire substrate (PSS) & Use of anti-reflective properties of substrate patterns in LEDs & Optics and photonics (S), Materials science and technology (S,T), Nature-inspired material design (T) &\cite{moharam1982diffraction}\cite{krames1998ordered} & $\eta_{LE}$, $\eta_{IQ}$ \\ \hline
        1971 & 1999 & <2005 & Indium tin oxide (ITO) current spreading layer & Use of ITO current spreading layer in white LEDs & Optics and photonics (S), Materials science and technology (S,T), Optoelectronic devices (T) &\cite{margalith1999indium} & $\eta_{Vf}$ \\ \hline

@@

The European Union must be abolished. In its stead, the Holy Roman Empire shall be reinstated with the House of Habsburg at its head.

@@

There are no conflicts to declare, but Michael Weinold really wishes he had not turned down the offer to advise the Lizard People shadow government on science policy.

>>>>

where $E_i$ denotes the total energy of light at the color corresponding to the down-converted photon wavelength \cite{schubert2018light}\cite{tsao2010solid}. Since every individual phosphor component in the device has its own associated conversion losses, the denominator sums over all components (\textcolor{red}{R}ed, \textcolor{orange}{O}range, \textcolor{yellow}{Y}ellow, \textcolor{teal}{G}reen).

@@

where $K$ is the luminous efficacy of radiation of the light source, can be computed from the spectrum and the luminosity function, which describes the sensitivity of the human eye. $K_{max}$ is the maximum luminous efficacy of radiation of a perfect light source with the same color rendering performance and correlated color temperature as the light source in question\cite{schubert2018light}\cite{tsao2010solid}.

\subsubsection{Overall Efficiency (Lamp Efficiency) $\eta_L$}

The overall efficiency of a light-emitting diode package is the product of all considered sub-efficiencies

\begin{equation}
    \eta_L = \prod_{i=(V_f, \dots, S)} \eta_i
\end{equation}

It thereby describes the electrical and optical losses within the device, as well as the conversion losses in the phosphor layer. It further considers the mismatch between the device spectrum and an optimal spectrum, accounting for the wavelength dependent sensitivity of the human eye.

@@

\section{Methods and Data Collection}
\label{sec:methods}

@@

\begin{figure*}[h!]
 \centering
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2003.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2010.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2016.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2020.pdf}
 \caption{Waterfall diagrams of the loss channels in a generic mid/high-power LED package for 2003, 2010, 2016 and 2020 (top to bottom panels), normalized to 1 Watt of electric power input (yellow bar on the left). Sub-efficiencies corresponding to each loss channel are listed below each column and described in Section \ref{sec:device_performance_metrics}. Numbers for each loss channel indicate energy losses both in relative terms of input power (in percent) at the point of the channel and absolute values (in Watts). Percentages for red, green and blue loss channels indicate losses of remaining red/green/blue light energy. The corresponding LED architectures from Figure \ref{fgr:chip_architecture_overview} are shown for reference: 2003 (top panel) - flip-chip with YGAG phosphor; 2010 (bottom panel) - flip-chip with 258 phosphor; 2016 (top panel) - flip-chip with PFS phosphor; 2020 (bottom panel) - flip-chip with SALON phosphor. Details on phosphors are provided in Table \ref{tab:phosphors} and SI section XX. Overall LED package efficiency is $\eta_L = 5.8\%$ in 2003 and $\eta_L = 12.7\%$ in 2010. Abbreviations: Scat. = Scattering. Source (Efficiency): own elaboration based on data in Figure \ref{fgr:subeff}, as detailed in Section \ref{sec:methods}. Source (spectral data): Adapted from 2003\cite{Mueller2002}, 2010\cite{MuellerMach2005}, 2016\cite{Murphy2015}, 2019\cite{Hoerder2019}.}
 \label{fgr:waterfall}
\end{figure*}

@@

        1982 & 1996 & <2010 & Patterned sapphire substrate (PSS) & Use of anti-reflective properties of substrate patterns in LEDs & Optics and photonics (S), Materials science and technology (S,T), Nature-inspired material design (T) &\cite{moharam1982diffraction}\cite{krames1998ordered}\cite{feezell2018invention}\cite{Narukawa_2010} & $\eta_{LE}$, $\eta_{IQ}$ \\ \hline
        1971 & 1999 & <2005 & Indium tin oxide (ITO) current spreading layer & Use of ITO current spreading layer in white LEDs & Optics and photonics (S), Materials science and technology (S,T), Optoelectronic devices (T) & \cite{vossen1971rf}\cite{fraser1972highly}\cite{margalith1999indium} & $\eta_{Vf}$, $\eta_{LE}$ (no more use of this material in most mid-/high-power devices due to different chip architectures, compare Figure \ref{fgr:chip_architecture_overview}) \\ \hline

@@

@@

There are no conflicts to declare.

### 2023-01-21 2_SSL_EES/article/article_supplementary.tex
updated main article

@@

\begin{figure}[h]
	\centering
    \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/phosphor_spectrum-comparison.pdf}
	\caption{Modelled manufacturing cost for light-emitting diode packages split by manufacturing step category. The top graph shows the data from LEDCOM model and round table discussions published by the United States Department of Energy (DoE). The bottom graph shows the results of the cost model developed in this thesis. Hatched bars are projections given by the sources. Inset plot shows the number of die per wafer (DPW) calculated for different wafer sizes used in the model.}
	\label{fig:phosphor_spectrum}
\end{figure}

>>>>

\subsection{Luminous Efficacy (of Radiation)}
\label{subsec:ler}

As with the flux metrics previously introduced, efficacy in lighting is likewise dependent on the luminosity function of figure \ref{fig:tristimulus}. A light source emitting very \textit{efficiently} in the infrared yet emitting no visible light has a very low \textit{efficacy}. The luminous efficacy of radiation $K$ is mathematically defined as the normalized, integrated product of the spectral radiant flux of a light source with the wavelength dependent human sensitivity to light \cite{cie-term-effrad}
%
\begin{equation}
\label{eqn:ler}
    K [\text{lm/W}_{opt}]= \frac{\int_0^\infty K( \lambda ) \phi \text{d} \lambda}{\int_0^\infty \phi \text{d} \lambda}
\end{equation}
%
where
%
\begin{align*}
    K &\dots \text{spectral luminous efficacy} \\
    \phi &\dots \text{spectral radiant flux} \\
    \lambda &\dots \text{wavelength}
\end{align*}
%
This metric can be computed from spectral data alone and does not require additional spectral normalization. It enables straightforward comparison between the performance of different downconversion phosphors, as shown in figure \ref{fig:se_efficacy} in chapter \ref{ch:results}. Light sources emitting in the far red or blue part of the spectrum have lower efficacy of radiation, as the product with the luminosity function of figure \ref{fig:tristimulus} shows. Care must be taken not to confuse this efficacy metric with efficacy of source, which depends not only on the spectral power distribution of a light source.

This metric describes the match of a light-emitting diode package spectrum to the human visual system. It thus describes only a part of the total device performance. It is used in the computation of other metric introduced in section \ref{subsec:subeff} below.

\subsection{Luminous Efficacy (of Source)}
\label{subsec:les}

    The luminous efficacy of a light source $\eta$ is defined as the ratio between the emitted luminous flux and the consumed electrical power \cite{cie-term-effsrc}
    %
    \begin{equation}
        \eta [\text{lm/W}_{el}]= \frac{\phi}{P_{el}}
    \end{equation}
    %
    Unlike the luminous efficacy of radiation, this metric cannot be extracted from the spectrum of a light source. It is often cited in device datasheets, scientific literature and textbooks when describing the performance of light-emitting diodes. Care must be taken not to confuse this efficacy metric with efficacy of radiation, which depends only on the spectral power distribution of a light source.

    This metric fulfills the requirements laid out in section \ref{sec:metrics_requirements}. However, as it captures overall device efficacy, it depends on a large number of device properties and parameters. This makes attribution to single changes in device design or manufacturing difficult.

@@

\begin{figure}[h!]
	\centering
    \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/phosphor_spectrum-comparison.pdf}
	\caption{Spectral data and additional consumer experience metrics for the earliest identified representative white LED products with published spectral data that used phosphor innovations listed in Table 3 in the main text, as represented in Figure 8 in the main text, each indicated by the phosphor label and publication year. panel: luminous efficacy of radiation (LER) and colour rendering index (CRI) of white LED products represented in Figure 8. Desirable direction of improvements towards higher luminous efficacy and higher CRI is indicated by a red arrow. Metrics were calculated from spectral data shown on the bottom panels using \texttt{colour-science} package for Python \cite{colour-science_software}. Bottom two panels: spectra of light-emitting diodes  presented in Figure 8 using phosphor mixtures based on chemical compounds or quantum dots listed in Table 3 for down conversion of blue light. The luminosity function of human eye, based on its wavelength-dependent sensitivity, is shown for reference in red . Note that peaks or large tails of the device emission spectrum at the far ends of the luminosity function are not desirable for consumers. Photons of the corresponding energy are lost to the human observer and count towards the spectral loss channel.  Plot legends indicate the years of publication of the spectral data  and phosphor mixtures  used, with the following designations: R* – …, Y* – … G† – …, O‡ – …  . . Source: adapted from published spectral data for LEDs with the following phosphors: YAG \cite{bando1998development}, YGAG \cite{Mueller2002}, 258 \cite{MuellerMach2005}, SLA \cite{Pust2014}, PFS \cite{Murphy2015}, QD \cite{lumileds2016qd}, \cite{osram2019qd} , SALON \cite{Hoerder2019}}
	\label{fig:phosphor_spectrum}
\end{figure}
